---
title: "RWorksheet_BIBIT#1"
output: pdf_document
date: "2024-09-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Set up a vector named age, consisting of 34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41,
51, 35, 24, 33, 41.
a. How many data points? 

    There are 34 Data Points

b. Write the R code and its output.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
         35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
         50, 37, 46, 25, 17, 37, 42, 53, 41, 51,
         35, 24, 33, 41)
length(age)
```
2. Find the reciprocal of the values for age.
    Write the R code and its output.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
         35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
         50, 37, 46, 25, 17, 37, 42, 53, 41, 51,
         35, 24, 33, 41)
reciprocal_age <- 1 / age
reciprocal_age
```
3. Assign also new_age <- c(age, 0, age).
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
         35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
         50, 37, 46, 25, 17, 37, 42, 53, 41, 51,
         35, 24, 33, 41)
new_age <- c(age, 0, age)

```
What happen to the new_age?

    The c() function is utilized to combine the vectors,
    resulting in the new_age vector, which comprises 69 elements:
    the initial 34 elements from the age vector, 
    followed by a single 0, and finally, 
    the duplicate 34 elements from the age vector.
    
\pagebreak

4. Sort the values for age.

    Write the R code and its output.
```{r}
sort(age)
```


5. Find the minimum and maximum value for age.
```{r}
min_age <- min(age)
max_age <- max(age)
print(paste("Minimum age:", min_age))
print(paste("Maximum age:", max_age))
```

6. Set up a vector named data, consisting of 2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5,
2.3, 2.5, 2.3, 2.4, and 2.7.


a. How many data points?

    12
      
b. Write the R code and its output.
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
length(data)
```


7. Generates a new vector for data where you double every value of the data. 
```{r}
data_doubled <- data * 2
print(data_doubled)
```


What happen to the data?


    We created a new vector data_doubled by 
    doubling every value in the original data vector, 
    resulting in a new vector with the same number of elements
    but with each element twice
    the value of the corresponding element in the original vector.
   
   \pagebreak
   
8.1 Integers from 1 to 100

```{r}
data_1_to_100 <- 1:100
data_1_to_100
print(data_1_to_100)
```



8.2 Numbers from 20 to 60
```{r}
data_20_to_60 <- 20:60
data_20_to_60
print(data_20_to_60)
```


8.3 Mean of numbers from 20 to 60
```{r}
mean_20_to_60 <- mean(20:60)
mean_20_to_60
print(mean_20_to_60)
```


8.4 Sum of numbers from 51 to 91
```{r}
sum_51_to_91 <- sum(51:91)
sum_51_to_91
print(sum_51_to_91)
```

\pagebreak

*8.5 Integers from 1 to 1,000
```{r}
data_1_to_1000 <- 1:1000
data_1_to_1000
```

    a. How many data points from 8.1 to 8.4?_______


    So, the total number of data points from 8.1 to 8.4 is: 100 + 41 + 1 + 1 = 143

    Therefore, the answer is: 143

    b. Write the R code and its output from 8.1 to 8.4.
    8.1 

```{r}
data_1_to_100 <- 1:100
data_1_to_100

```

\pagebreak

    8.2
```{r}
data_20_to_60 <- 20:60
data_20_to_60

```

    8.3
```{r}
mean_20_to_60 <- mean(20:60)
mean_20_to_60

```

    8.4
```{r}
sum_51_to_91 <- sum(51:91)
sum_51_to_91

```    
    
    c. For 8.5 find only maximum data points until 10.
```{r}
data_1_to_1000 <- 1:1000
data_1_to_10 <- data_1_to_1000[1:10]
print(data_1_to_10)
```
    
    9. *Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and 7 using filter option.
    
```{r}
vector <- Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(1, 100, 1))
print(vector)
```
  
    10. Generate a sequence backwards of the integers from 1 to 100.
    Write the R code and its output.
```{r}
vector <- seq(100, 1, -1)
print(vector)
```
    
    11. List all the natural numbers below 25 that are multiples of 3 or 5.

```{r}
multiples <- seq(1, 24, 1)[(seq(1, 24, 1) %% 3 == 0) | (seq(1, 24, 1) %% 5 == 0)]
print(multiples)
```

    Find the sum of these multiples.
```{r}
sum_multiples <- sum(multiples)
print(sum_multiples)
```
    a. How many data points from 10 to 11?
    There are 112 data points
    
    b. Write the R code and its output from 10 and 11.
10
```{r}
vector <- seq(100, 1, -1)
print(vector)
```
11
```{r}
multiples <- seq(1, 24, 1)[(seq(1, 24, 1) %% 3 == 0) | (seq(1, 24, 1) %% 5 == 0)]
print(multiples)
```
      
      12. Statements can be grouped together using braces ‘{’ and ‘}’.
      A group of statements is sometimes called a block. 
      Single statements are evaluated when a new line is 
      typed at the end of the syntactically complete statement. 
      Blocks are not evaluated until a new line is 
      entered after the closing brace.

      Enter this statement:
      x <- {0 + x + 5 + }

    Describe the output.
    
    When you enter x <- {0 + x + 5 + }, 
    you'll get an error because the statement is incomplete
    and R expects more input. 
    The trailing + symbol indicates R is waiting for another operand,
    and the { brace isn't properly closed, causing the error.
    
  \pagebreak
    
    13. *Set up a vector named score, 
    consisting of 72, 86, 92, 63, 88, 89, 91, 92, 75, 75 and 77.
    To access individual elements of an atomic vector,
    one generally uses the x[i] construction.
    Find x[2] and x[3]. 
    
    Write the R code and its output.
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)
score[2]
score[3]
```
    
    
    14. *Create a vector a = c(1,2,NA,4,NA,6,7).
```{r}
a <- c(1, 2, NA, 4, NA, 6, 7)
```
    
a. Change the NA to 999 using the codes print(a,na.print="-999").
```{r}
a <- c(1, 2, NA, 4, NA, 6, 7)
print(a, na.print="-999")
```

b. Write the R code and its output. Describe the output.
```{r}
a <- c(1, 2, NA, 4, NA, 6, 7)
print(a, na.print="-999")
```

    The output shows the vector a with the NA values replaced with -999 when printed.
    The actual values in the vector a remain unchanged,
    only the way they are printed is affected by the na.print argument.
 
    15. A special type of function calls can appear on the left hand side of the assignment
    operator as in > class(x) <- "foo".   
    
```{r}
  name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
```






