---
title: "RWorksheet_BIBIT#2"
output:
  pdf_document: default
  html_document: default
date: "2024-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using Vectors
  1. Create a vector using: operator
    a. Sequence from -5 to 5. Write the R code and its output. 
```{r}
        
        x <- -5:5
        x
    
```
    
        b.  x <- 1:7. What will be the value of x?
    
```{r}
    
        x <- 1:7
        x
    
```


  2.* Create a vector using sec() function
    a.seq(1,3 by =0.2) # specify step size 
    Write R Script and its output. describe the output
```{r}
    
        a.seq <- seq(1, 3, by = 0.2)
        a.seq
    
```

        The seq function in R is used to generate a sequence of   numbers. 
        In this case, we are generating a sequence from 1 to 3    with a step size of 0.2.
        The output is a vector of 11 numbers, starting from 1 and incrementing by 0.2 up to 3.

\pagebreak

    3. A factory has a census of its workers. 
      There are 50 workers in total. 
      The following list shows their ages:
      
      34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
      50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41,
      48, 27, 39, 19, 30, 61, 54, 58, 26, 18. 
      

a. Access 3rd element, what is the value?

```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
          50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41,
          48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
ages[3]
```



b. Access 2nd and 4th element, what are the values?
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
          50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41,
          48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
ages[2]
ages[4]
```



e. Access all but the 4th and 12th element is not included. Write the R script and its output.

```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49,
          50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41,
          48, 27, 39, 19, 30, 61, 54, 58, 26, 18)


exclude_indices <- c(4, 12)


ages[-exclude_indices]
```

4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then     named the vector, names(x).

  a. Print the results. Then access x[c("first", "third")].
  Describe the output.
```{r}

x <- c("first"=3, "second"=0, "third"=9)

print(x)

print(x[c("first", "third")])
```


  b. Write the code and its output..
  ```{r}

x <- c("first"=3, "second"=0, "third"=9)

names(x)
  ```
  
5. Create a sequence x from -3:2.

a. Modify 2nd element and change it to 0; x[2] <- 0 X Describe the output.

```{r}

x <- -3:2


print(x)


x[2] <- 0


print(x)
```

b. Write the code and its output.
```{r}


x <- -3:2

x[2] <- 0

x
```

6. *The following data shows the diesel fuel purchased by Mr. Cruz.

Month Jan Feb March Apr May June
Price per liter (PhP) 52.50 57.25 60.00 65.00 74.25 54.00
Purchase–quantity(Liters) 25 30 40 50 10 45

a. Create a data frame for month, price per liter (php)
and purchase-quantity (liter). Write the R scripts and
its output.

```{r}

df <- data.frame(
  Month = c("Jan", "Feb", "March", "Apr", "May", "June"),
  Price_per_liter = c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00),
  Purchase_quantity = c(25, 30, 40, 50, 10, 45)
)

print(df)
```

\pagebreak

b. What is the average fuel expenditure of Mr. Cruz from
Jan to June? Note: Use ‘weighted.mean(liter,
purchase)‘. Write the R scripts and its output.

```{r}

df <- data.frame(
  Month = c("Jan", "Feb", "March", "Apr", "May", "June"),
  Price_per_liter = c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00),
  Purchase_quantity = c(25, 30, 40, 50, 10, 45)
)


average_expenditure <- weighted.mean(df$Price_per_liter, df$Purchase_quantity)


print(average_expenditure)


```
  
  
7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.

a. Type “rivers” in your R console.
Create a vector data with 7
elements, containing the number of elements (length)
in rivers, their sum (sum), mean (mean),
median(median), variance(var), standard deviation(sd),
minimum (min) and maximum (max).

```{r}
data(rivers)

vector_data <- c(
  length = length(rivers),
  sum = sum(rivers),
  mean = mean(rivers),
  median = median(rivers),
  variance = var(rivers),
  sd = sd(rivers),
  min = min(rivers),
  max = max(rivers)
)


print(vector_data)
```

data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),
sd(rivers), min(rivers), max(rivers))

\pagebreak

b. What are the results?

The results of the "rivers" dataset analysis are:

length(rivers): The number of elements in the "rivers" dataset, which is 141.

sum(rivers): The sum of the lengths of the rivers, which is 7442.

mean(rivers): The mean length of the rivers, which is approximately 52.86 miles.

median(rivers): The median length of the rivers, which is 44 miles.

var(rivers): The variance of the lengths of the rivers, which is approximately 441.17.

sd(rivers): The standard deviation of the lengths of the rivers, which is approximately 21.36 miles.

min(rivers): The minimum length of the rivers, which is 10 miles.

max(rivers): The maximum length of the rivers, which is 3710 miles.


c. Write the R scripts and its outputs.
```{r}

data(rivers)

data <- c(
  length(rivers),
  sum(rivers),
  mean(rivers),
  median(rivers),
  var(rivers),
  sd(rivers),
  min(rivers),
  max(rivers)
)


print(data)
```

\pagebreak

The table below gives the 25 most powerful celebrities and their annual pay as ranked
by the editions of Forbes magazine and as listed on the Forbes.com website.

a. Create vectors according to the above table.
Write the R scripts and its output.

```{r}
# a. Create vectors for celebrity names, power rankings, and pay
celebrity_names <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", 
                     "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", 
                     "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
                     "Paul McCartney", "George Lucas", "Elton John", "David Letterman",
                     "Phil Mickelson", "J.K Rowling", "Bradd Pitt", "Peter Jackson", 
                     "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")

power_ranking <- c(1:13, 14:25)

pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 
         40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)


celebrity_names
power_ranking
pay


```

\pagebreak

b. Modify the power ranking and pay of J.K. Rowling.
Change power ranking to 15 and pay to 90. Write the
R scripts and its output.

```{r}


power_ranking[power_ranking == 19] <- 15 
pay[celebrity_names == "J.K Rowling"] <- 90


celebrity_names
power_ranking
pay

```

\pagebreak

c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?

```{r}

PowerRanking <- data.frame(
  PowerRanking = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25),
  CelebrityName = c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
"Steven Spielberg", "Howard Stern", "50 Cent","Cast of the Sopranos", "Dan Brown", "Bruce Springsteen", 
"Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman",
"Phil Mickelson", "J.K Rowling", "Bradd Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant"),
  Pay = c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)
)


write.csv(PowerRanking, "PowerRanking.csv", row.names = FALSE)


PowerRanking <- read.csv("PowerRanking.csv")


print(PowerRanking)

```

\pagebreak

d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.

```{r}
power_ranking <- c(1:13, 14:25)
celebrity_name <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", 
                   "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", 
                   "Cast of the Sopranos", "Dan Brown", "Bruce Springsteen", 
                   "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", 
                   "Elton John", "David Letterman", "Phil Mickelson", "J.K Rowling", 
                   "Bradd Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon",
                   "Celine Dion", "Kobe Bryant")
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34,
         40, 47, 75, 25, 39, 45, 32, 40, 31)
Forbes_Ranking <- data.frame(power_ranking, celebrity_name, pay)
Forbes_Ranking
Ranks <- Forbes_Ranking[10:20, ]
Ranks

save(Ranks, file = "Ranks.RData")
```

e. Describe its output.

The output of the R script is a data frame named Ranks that contains rows 10 to 20 of the original data frame Forbes_Ranking. The data frame Ranks has 11 rows and 3 columns.

Here is the output:

```{r}
Ranks
```

9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
a. Import the excel file into your RStudio.
What is the R script?

```{r}
library(readxl)
hotels_vienna <- read_excel("/cloud/project/rigor/worksheet2/hotels-vienna.xlsx")

```

b. How many dimensions does the dataset have?
What is the R script? WHat is its output?

```{r}

dim(hotels_vienna)

```


c. Select columns country, neighbourhood,
price, stars, accomodation_type, and
ratings. Write the R script.

```{r}


library(dplyr)
hotels_vienna_selected <- hotels_vienna %>%
  select(country, neighbourhood, price, stars, accommodation_type, rating)

```

\pagebreak


d. Save the data as **new.RData to your RStudio. Write the
R script.


```{r}
save(hotels_vienna_selected, file = "vienna-hotel-selected.RData")
```


e. Display the first six rows and last six rows of the
new.RData. What is the R script?

```{r}

load("vienna-hotel-selected.RData")


head(hotels_vienna_selected, 6)


tail(hotels_vienna_selected, 6)
```

10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down.

a. Write the R scripts and its output.
```{r}
# Create a list of 10 vegetables
vegetables <- list("Kalabasa", "Kapayas", "Monggo", "Kangkong", "Okra", "Malunggay", "Potatoes", "Ampalaya", "Cabbage", "Asparagus")

# Print the list of vegetables
vegetables
```


b. Add 2 additional vegetables after the last vegetables
in the list. What is the R script and its output?

```{r}

vegetables <- c(vegetables, "Tomatoes", "Eggplant")


vegetables
```


c. Add 4 additional vegetables after index 5. How many
datapoints does your vegetable list have? What is the
R script and its output?

```{r}


vegetables <- c(vegetables[1:5], list("Pechay", "Bawang", "Sitaw", "Singkamas"), vegetables[6:12])

vegetables
```


d. Remove the vegetables in index 5, 10, and 15. How
many vegetables were left? Write the codes and its
output.

```{r}

vegetables <- vegetables[-c(5, 10, 15)]

vegetables
```


Note: Do not forget to push into your GitHub repo.

Prepared by:
JOYCE F. JAMILE
Instructor
























