---
title: "RWorksheet_BIBIT#4b"
author: "John Benedict Bibit"
date: "2024-10-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r}

vectorA <- c(1, 2, 3, 4, 5)


matrix_result <- matrix(0, nrow = 5, ncol = 5)


for (i in 1:length(vectorA)) {
  matrix_result[1, i] <- vectorA[i]
}


print(matrix_result)
```



2. Print the string "*" using for() function. The output should be the same as shown
in Figure
```{r}

n <- 5


for (i in 1:n) {

  cat(rep("*", i), "\n")
}
```

\pagebreak 
3.Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
start_num <- 5  # Assigned starting number for the Fibonacci sequence
a <- 0
b <- 1

cat("Fibonacci sequence starting from", start_num, "up to 500:\n")

repeat {
  fib <- a + b
  a <- b
  b <- fib
  
  if (fib >= start_num) {
    cat(fib, "\n")
  }
  
  if (fib > 500) {
    break
  }
}
  
```



4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result

```{r}

shoes_data_csv <- read.csv("/cloud/project/rigor/worksheet4/shoes_data.csv")


head(shoes_data_csv)
```



b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.

```{r}


shoes_data_csv

male_data <- subset(shoes_data_csv, Gender == "M")
female_data <- subset(shoes_data_csv, Gender == "F")

num_male <- nrow(male_data)
num_female <- nrow(female_data)

cat("Number of observations in Male:", num_male, "\n")
cat("Number of observations in Female:", num_female, "\n")
```


c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.

```{r}
library(readr)

household_data <- read.csv("/cloud/project/rigor/worksheet4/CopyOfHouseholdData.csv")

gender_count <- table((household_data$Sex))

if (length(gender_count) > 0) {
    bar_colors <- c("blue", "pink")
    
    barplot(gender_count,
            main = "Number of Males and Females in Household Data",
            xlab = "Gender",
            ylab = "Number of Observations",
            col = bar_colors,
            beside = TRUE,
            names.arg = names(gender_count))
    
    legend("topright", legend = names(gender_count), fill = bar_colors)
} else {
    print("No valid gender data to plot.")
}
```


\pagebreak

5. The monthly income of Dela Cruz family was spent on the following:
Food Electricity Savings Miscellaneous
60 10 5 25


a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
values <- c(60, 10, 5, 25)

percentages <- values / sum(values) * 100

pie(percentages, 
    labels = paste(categories, round(percentages, 1), "%"), 
    col = c("lightblue", "lightgreen", "lightpink", "lightyellow"),
    main = "Dela Cruz Family Monthly Income Distribution")

legend("topright", legend = categories, fill = c("lightblue", "lightgreen", "lightpink", "lightyellow"))
```

