---
title: "RWorksheet_BIBIT#3b"
author: "John Benedict Bibit"
date: "2024-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create a data frame using the table below.

a. Write the codes.

```{r}
respondents_data <- data.frame(
  Respondents = 1:20,
  Sex = c(2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2),
  Fathers_Occupation = c(1, 3, 3, 3, 1, 2, 3, 1, 1, 1, 3, 2, 1, 3, 3, 1, 3, 1, 2, 1),
  Persons_at_Home = c(5, 7, 3, 8, 5, 9, 6, 7, 8, 4, 7, 5, 4, 7, 8, 8, 3, 11, 7, 6),
  Siblings_at_School = c(6, 4, 4, 1, 2, 1, 5, 3, 1, 2, 3, 2, 5, 5, 2, 1, 2, 5, 3, 2),
  Types_of_Houses = c(1, 2, 3, 1, 1, 3, 3, 1, 2, 3, 2, 3, 2, 2, 3, 3, 3, 3, 3, 2)
)
```

b. Describe the data. Get the structure or the summary of the data
```{r}
str(respondents_data)
summary(respondents_data)
```

\pagebreak

c. Is the mean number of siblings attending is 5?
```{r}
mean_siblings <- mean(respondents_data$Siblings_at_School)
print(paste("Mean number of siblings attending school:", mean_siblings))
print(paste("Is the mean number of siblings attending 5?", mean_siblings == 5))
```


d. Extract the 1st two rows and then all the columns using the subsetting functions.
Write the codes and its output.
```{r}
first_two_rows <- respondents_data[1:2, ]
print(first_two_rows)
```

e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
selected_rows_cols <- respondents_data[c(3,5), c(2,4)]
print(selected_rows_cols)
```


f. Select the variable types of houses then store the vector that results as types_houses.
Write the codes.
```{r}
types_houses <- respondents_data$Types_of_Houses
print(types_houses)
```


g. Select only all Males respondent that their father occupation was farmer. Write the
codes and its output.
```{r}
male_farmer <- respondents_data[respondents_data$Sex == 1 & respondents_data$Fathers_Occupation == 1, ]
print(male_farmer)
```
\pagebreak

h. Select only all females respondent that have greater than or equal to 5 number of
siblings attending school. Write the codes and its outputs.
```{r}
female_5siblings <- respondents_data[respondents_data$Sex == 2 & respondents_data$Siblings_at_School >= 5, ]
print(female_5siblings)
```


2. Write a R program to create an empty data frame. Using the following codes:
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
a. Describe the results.
```{r}
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
```


3. Create a .csv file of this. Save it as HouseholdData.csv
```{r}
write.csv(respondents_data, "HouseholdData.csv", row.names = FALSE)
```

a. Import the csv file into the R environment. Write the codes.
```{r}
imported_data <- read.csv("HouseholdData.csv")
```

\pagebreak

b. Convert the Sex into factor using factor() function and change it into integer.[Legend:
Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
imported_data$Sex <- factor(imported_data$Sex, levels = c("Male", "Female"))
```


c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood
= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r}
imported_data$Types_of_Houses <- factor(imported_data$Types_of_Houses, 
                                        levels = c("Wood", "Semi-Concrete", "Concrete"))

```


d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What
is the R code and its output?
```{r}
imported_data$Fathers_Occupation <- factor(imported_data$Fathers_Occupation, 
                                           levels = c("Farmer", "Driver", "Others"))
```

e. Select only all females respondent that has a father whose occupation is driver. Write
the codes and its output.
```{r}
female_driver_fathers <- respondents_data[respondents_data$Sex == "Female" & respondents_data$Fathers_Occupation == "Driver", ]
print(female_driver_fathers)
```


f. Select the respondents that have greater than or equal to 5 number of siblings attending
school. Write the codes and its output.
```{r}
five_or_more_siblings <- respondents_data[respondents_data$Siblings_at_School >= 5, ]
print(five_or_more_siblings)
```


4. Interpret the graph.


The sentiment analysis graph is a stacked bar chart that categorizes public sentiment across five topics: Technology, Food, Politics, Sports, and Entertainment. It shows that Technology has the highest positive sentiment, while Politics has the most negative sentiment, reflecting public frustration. Food has a balanced distribution of positive, negative, and significant neutral sentiments, indicating mixed feelings. Sports display slightly more positive than negative sentiments, and Entertainment leans towards positivity with nearly 50% positive reactions. Overall, the graph highlights varying sentiments across these categories, providing insights into public opinion that can inform marketing strategies and content creation.



